\frontmatter

\maketitle
\thispagestyle{empty}

\newpage
\tableofcontents

\newpage
\chapter*{Preface to the First Edition}
\addcontentsline{toc}{chapter}{Preface to the First Edition}
\markboth{Preface}{Preface}

This short excerpt is meant as a how-to-use this book guide or a "where to find things" guide. Every chapter contains content which helps explain concepts behind mathematical proofs. I have also tried to choose some examples which are commonly used in competition mathematics, with the goal of proving these ideas. Each chapter also has exercises, to which solutions are in Appendix B. Appendix B also contains solutions to the end-of-chapter problem set. Try not to rely too much on the solutions!

Each problem is also marked with a $\star$ if I believe it is an especially difficult problem. I'm not perfect though! Difficulty is also a subjective thing; different people find different things to be difficult. The take-away is that regardless of the status of the problem, do not be discouraged if a problem takes a long time or is especially difficult. The struggle is where learning happens! To help guide you, I have also included a Hints appendix (Appendix A). Problems with hints have them labeled in the order they are to be read. For example, for Problem 2.6, you will want to read Hint H71 before you read hint H57 (as is listed in the order H71, H57). The hints are scrambled so looking at one hint doesn't accidentally reveal the next hint. The most important advice for solving the problems at the end of the chapter is: do not give up!

I have also assumed familiarity with some basic concepts that are typically not taught in school. The major topic is Modular Arithmetic. If you are not too familiar, don't worry! Appendix C (Additional Resources) has a section devoted to this with some good handouts that teach Modular Arithmetic. If things still do not make sense, then use your resources! I have attached some resources I have found useful in my math journey, but if none are satisfactory then there do exist search engines and Generative AI tools to help (though I would warn to use them cautiously). 

This is an independent project to help bring more awareness and knowledge about mathematical proofs, which I often find to be lacking in most educational contexts until college. A typical math major encounters his/her first proof during sophomore year of college, or later. This book aims to bring that wait-time down drastically; math is all about proofs so it does not make sense to wait for 20 or so years before learning about it!

I hope this book serves as a guide, an introduction, and an inspiration to learn about, study, and enjoy mathematics. 

\bigskip
\noindent
\textit{Hrishikesh Shanmuganathan - Author}